Frozen Civil Conflicts

### 2024-03-07 paper.qmd
Draft edits

Finally, it is worth distinguishing frozen conflict from several related concepts, specifically protracted stalemates and enduring internal rivalries. In the case of the former, stalemates during a conflict can potentially endure for extended periods of time and offer conditions similar to those of a frozen conflict. However, they do not necessitate a cessation of active fighting. Instead, they may develop into wars of attrition where actors are unable to achieve an outright military advantage, but continue fighting in the hope of outlasting and "grinding down" the enemy, *i.e.* being better able to endure the prolonged costs of conflict. In the case of the latter, enduring internal rivalries are defined by @derouen_enduring_2008 as "...internal conflicts between a government and an insurgency with at least 10 years of [non-consecutive] armed conflict in which there are at least 25 deaths", an extension of the interstate concept of enduring rivalries [@goertz_enduring_1993] characterized by repeated onsets of militarized disputes. Compared to frozen conflicts, enduring internal rivalries include an explicit temporal requirement and, similar to stalemates, are not restricted to conflicts where fighting has largely stopped. The concept of frozen concepts uniquely captures the post-war condition of unstable peace that characterizes certain conflicts where there is no conclusive termination to the initial episode of fighting.

## How do frozen conflicts emerge?

Frozen conflict as an equilibrium of a situation where both actors strictly prefer a military victory but are unable to achieve a significant military advantage and are unwilling to acquiesce to a peace agreement. The belligerents will seek to minimize costs by de-escalating the conflict intensity and the conflict will eventually freeze.

@@

-  transitory war gains outweigh peaceful division of contested good
- time horizon, low discount factor (Stall war paper assumes that actors actually discount the future?). Anticipation of a future favourable shift in the balance of power
- credible commitment issues undermine payoff from a peace agreement

Unlike conflict onsets, assume that fighting has revealed information about actors' resolve and capabilities.

>>>>

Finally, it is worth distinguishing frozen conflict from several related concepts, specifically protracted stalemates and enduring internal rivalries. In the case of the former, stalemates during a conflict can potentially endure for extended periods of time and offer conditions similar to those of a frozen conflict. However, they do not necessitate stop active fighting. Instead, they may develop into wars of attrition where actors are unable to achieve an outright military advantage, but continue fighting in the hope of outlasting and "grinding down" the enemy, *i.e.* being better able to endure the prolonged costs of conflict. In the case of the latter, enduring internal rivalries are defined by @derouen_enduring_2008 as "...internal conflicts between a government and an insurgency with at least 10 years of [non-consecutive] armed conflict in which there are at least 25 deaths", an extension of the interstate concept of enduring rivalries [@goertz_enduring_1993] characterized by repeated onsets of militarized disputes. Compared to frozen conflicts, enduring internal rivalries include an explicit temporal requirement and, similar to stalemates, are not restricted to conflicts where fighting has largely stopped. The concept of frozen concepts uniquely captures the post-war condition of unstable peace that characterizes certain conflicts where there is no conclusive termination to the initial episode of fighting.

## The Emergence of Frozen Conflicts

I view the emergence of frozen conflicts as an equilibrium state of a situation where actors strictly prefer a military victory, but are unable to achieve a significant military advantage and are unwilling to acquiesce to a comprehensive peace agreement. Faced with a stalemate, belligerents will seek to minimize costs by de-escalating the conflict intensity whilst maintaining a minimum degree of hostility. This may be done informally or formally --- the latter entailing ceasefires or partial peace agreements --- resulting in a cessation of large-scale active fighting following which the conflict will freeze.

To achieve this equilibrium state requires firstly a convergence in beliefs over the utility of additional fighting. Actors must believe that there is a high probability that additional fighting will simply maintain the status quo, which will decrease the incentive towards continued violence.[^convergence] From the previous definition, frozen conflict is a *post-conflict* condition that follows after the termination of the initial violent conflict. Thus, a stalemate can be considered a minimum necessary condition for the emergence of a frozen conflict if it incentivizes a conflict termination.[^stalemates]

[^stalemates]: Notably, I do not define stalemates strictly in terms of power parity since there are a number of reasons for which an actor may be unable to achieve a military advantage during the course of a conflict. This can, of course, include a balance in the material distribution of power, but also constraints arising from, for example, domestic audience costs and political interests of external, third party actors.

[^convergence]: This is akin to the "Principle of Converge" laid forth by @slantchev_principle_2003 whereby the usefulness of fighting decreases when it ceases to convey information on an opponent's capabilities and resolve

@@
